### 2022-05-07 final_manuscript_no_markup.tex
the real submitted version :tada: :rocket:
some editing for final version

% Title must be 250 characters or less.
\begin{flushleft}
{\Large
\textbf\newline{Ten simple rules to host an inclusive conference} % Please use "sentence case" for title and headings (capitalize only the first word in a title (or heading), the first word in a subtitle (or subheading), and any proper nouns).
}
\newline
% Insert author names, affiliations and corresponding author email (do not include titles, positions, or degrees).
\\
Rocío Joo\textsuperscript{1*\Yinyang}, 
Andrea Sánchez-Tapia\textsuperscript{2\Yinyang},
Sara Mortara\textsuperscript{3},
Yanina Bellini Saibene\textsuperscript{4,5},
Heather Turner\textsuperscript{6}, %\ddag},
Dorothea Hug Peter\textsuperscript{7}, %\ddag},
Natalia Soledad Morandeira\textsuperscript{8},
Matt Bannert\textsuperscript{9},
Batool Almazrouq\textsuperscript{10},
Elizabeth Hare\textsuperscript{11},
Laura Ación\textsuperscript{5,12},
Juan Pablo Narváez-Gómez\textsuperscript{13},
Marcela Alfaro Córdoba\textsuperscript{14},
Federico Marini\textsuperscript{15},
Rita Giordano\textsuperscript{16},
Silvia Canelón\textsuperscript{17},
Anicet Ebou\textsuperscript{18},
Adithi Upadhya\textsuperscript{19},
Joselyn Chávez\textsuperscript{20},
Janani Ravi\textsuperscript{21*}
%,3\textcurrency}
\\
\bigskip
\textbf{1} Global Fishing Watch, Washington, DC 20036, USA
\\
\textbf{2} Instituto de Pesquisas Jardim Botânico do Rio de Janeiro, Rio de Janeiro, Rio de Janeiro, Brazil
\\
\textbf{3} International Institute for Sustainability, Rio de Janeiro, Rio de Janeiro, Brazil
\\
\textbf{4} Instituto Nacional de Tecnología Agropecuaria, Ruta Nac. N° 5, Km. 580, Anguil, La Pampa, Argentina; Universidad Nacional Guillermo Brown, Buenos Aires, Argentina
\\
\textbf{5} MetaDocencia, Argentina
\\
\textbf{6} Department of Statistics, University of Warwick, Coventry, UK; The R Foundation
\\
\textbf{7} Swiss Federal Institute for Forest, Snow and Landscape Research WSL, Birmensdorf, Switzerland
\\
\textbf{8} Instituto de Investigación e Ingeniería Ambiental, Universidad Nacional de San Martín - Consejo Nacional de Investigaciones Científicas y Técnicas, General San Martín, Buenos Aires, Argentina
\\
\textbf{9} KOF Swiss Economic Institute, ETH Zurich, Zurich, Switzerland
\\
\textbf{10} Institute of Systems, Molecular and Integrative Biology, University of Liverpool, Liverpool, L69 7ZB, UK; King Abdullah International Medical Research Center, Riyadh, Saudi Arabia
\\
\textbf{11} Dog Genetics LLC, Astoria, NY, USA
\\
\textbf{12} Instituto de Cálculo, FCEyN, UBA-CONICET, Buenos Aires, Argentina
\\
\textbf{13} Departamento de Botânica, Instituto de Biociências, Universidade de São Paulo, São Paulo, São Paulo, Brasil
\\
\textbf{14} Department of Statistics, University of California, Santa Cruz, CA, USA
\\
\textbf{15} Institute of Medical Biostatistics, Epidemiology and Informatics (IMBEI), University Medical Center Mainz, Mainz, Germany
\\
\textbf{16} Royal Society of Chemistry, Thomas Graham House (290), Science Park, Milton Road, Cambridge, UK
\\
\textbf{17} Department of Biostatistics, Epidemiology \& Informatics, University of Pennsylvania, Philadelphia, PA, USA.
\\
\textbf{18} Bioinformatics team, Département de Formation et de Recherches en Agriculture et Ressources Animales, Institut National Polytechnique Félix Houphouët-Boigny, BP 1313 Yamoussoukro, Côte d’Ivoire.
\\
\textbf{19} ILK Labs, Bengaluru, Karnataka, 560046, India
\\
\textbf{20} Instituto de Biotecnología, UNAM, Cuernavaca, Morelos, México
\\
\textbf{21} Departments of Pathobiology and Diagnostic Investigation, Microbiology and Molecular Genetics, Michigan State University, East Lansing, MI 48824, USA
\\
\bigskip

@@

% Use the asterisk to denote corresponding authorship and provide email address in note below.
*Co-corresponding authors: rocio.joo@globalfishingwatch.org, janani@msu.edu

@@

Conferences are spaces to meet and to network within and across academic and technical fields, to learn about new advances, and share our work. They can help define career paths and create long-lasting collaborations and opportunities. 
However, these opportunities are not equal for all. 
This article introduces ten simple rules to host an inclusive conference based on the authors' recent experience organizing the 2021 edition of the useR! statistical computing conference, which attracted a broad range of participants from academia, industry, government, and the non-profit sector. 
Coming from different backgrounds, career stages, and even continents, we embraced the challenge of organizing a high-quality virtual conference in the context of the COVID-19 pandemic and making it a kind, inclusive, and accessible experience for as many people as possible.
The rules result from our lessons learned before, during, and after the organization of the conference. 
They have been written mainly for potential organizers and selection committees of conferences and contain multiple practical tips to help a variety of events become more accessible and inclusive. We see this as a starting point for conversations and efforts towards building more inclusive conferences across the world.
* Translated versions of the English abstract and the list of rules are available in ten languages in the Supporting Information: Arabic, French, German, Italian, Japanese, Korean, Portuguese, Spanish, Tamil, and Thai.

@@

% Roadmap of the paper
Here, we present a set of ten rules based on the lessons we learned before, during, and after the organization of the useR! conference.
\linelabel{methods-a} The rules were first drafted by the core team and the members of the diversity, accessibility, and inclusion team of useR! 2021 who wrote their own ten simple rules (see Supporting Information) based on the collective experience of organizing the conference and fueled by personal experiences and literature review.
A final list was proposed and then discussed and reviewed by the rest of the coauthors. \linelabel{methods-b}
The rules are organized in three sections: three foundation rules, six design rules, and a continuity rule (\textbf{Fig. \ref{fig:diagram}}).
The \textbf{foundation rules} comprise key elements to conceive the work on diversity and inclusion in any conference. 
\textbf{Rule 1} is about setting a vision of diversity and inclusion that should guide all the efforts and decision making in the organization.
\textbf{Rule 2} focuses on how to create a safe and welcoming environment for all the attendees. 
\textbf{Rule 3} highlights the importance of starting with an inclusive and diverse organizing team and provides tips on work dynamics.
The \textbf{design rules} focus on weaving inclusion into the conference design process.
In \textbf{Rule 4}, we introduce multiple ways to counteract bias in the conference program (keynotes, program committee, abstract selection, and thematic sessions). 
\textbf{Rule 5} provides advice for designing an inclusive online component in virtual and hybrid conferences.
\textbf{Rule 6} focuses on accessibility practices to include people with disabilities.
In \textbf{Rule 7}, we provide suggestions to account for linguistic diversity. 
\textbf{Rule 8} offers tips for developing an inclusive communication strategy. 
In \textbf{Rule 9}, we address budgeting for inclusive practices and helping participants with affordable registration costs, scholarships, and other forms of financial support.
Finally, 
\textbf{Rule 10}, the \textbf{continuity} rule, emphasizes the importance of self-assessment and advocates for making the conference part of a long-term commitment to inclusion and for passing the torch to future organizers.

@@

\section*{Acknowledgments}
The authors of this piece would like to thank every single member of the organizing team of useR! 2021 (\url{https://user2021.r-project.org/about/global-team}) for their valuable contribution to an inclusive conference experience and the R Foundation for trusting us with the organization of useR! 2021 and supporting us through the process. Special thanks to Saranjeet Kaur and gwynn sturdevant for suggestions to previous versions of the manuscript, and Francisco Etchart for his help with the Marmot figures. We would also like to thank Arjun Krishnan, Julia Ganz, Ingo Braasch, Susan Ewart, and Hilda Mejia-Abreu for their feedback on the manuscript. In addition to some of the authors (listed in the Translated Abstracts section of Supporting Information), we would like to acknowledge the following contributors for translating our extended abstract: Koki Tsuyuzaki and Kozo Nishida [Japanese], KwangChun Lee [Korean], Arjun Krishnan [Tamil], Kewalin Samart, Natnaree Yolnava, and Wanprakai Kaewphrae [Thai].
The comments of the anonymous reviewers contributed to the improvement of the manuscript and we thank them for that.

@@

JR is funded by start-up award from Michigan State University (MSU); publication costs were waived because of MSU’s institutional partnership with PLoS. HT was supported by the Engineering and Physical Sciences Research Council [EPSRC EP/V052128/1] during the preparation of this article.

The organization of useR! 2021---which inspired this work---was largely a volunteer effort (from many people including all the authors), but stipends were given from the conference funds for some specific roles.
AE, AU, AST, BA, JC, JPNG, MAC, NM, RG, RJ, SM, and YBS received stipends for roles that include code of conduct team, zoom host, and diversity and social program leads.

@@

\section*{Supporting Information}

Supporting Information 1. Abstract and list of rules in multiple languages: Arabic, French, German, Italian, Japanese, Korean, Portuguese, Spanish, Tamil, and Thai.

Supporting Information 2. Rules proposed in the process of shaping these ten simple rules.

>>>>

% Title must be 250 characters or less.
\begin{flushleft}
{\Large
\textbf\newline{Ten simple rules to host an inclusive conference} % Please use "sentence case" for title and headings (capitalize only the first word in a title (or heading), the first word in a subtitle (or subheading), and any proper nouns).
}
\newline
% Insert author names, affiliations and corresponding author email (do not include titles, positions, or degrees).
\\
Rocío Joo\textsuperscript{1*\Yinyang}, 
Andrea Sánchez-Tapia\textsuperscript{2\Yinyang},
Sara Mortara\textsuperscript{3},
Yanina Bellini Saibene\textsuperscript{4,5},
Heather Turner\textsuperscript{6}, %\ddag},
Dorothea Hug Peter\textsuperscript{7}, %\ddag},
Natalia Soledad Morandeira\textsuperscript{8},
Matt Bannert\textsuperscript{9},
Batool Almazrouq\textsuperscript{10},
Elizabeth Hare\textsuperscript{11},
Laura Ación\textsuperscript{5,12},
Juan Pablo Narváez-Gómez\textsuperscript{13},
Marcela Alfaro Córdoba\textsuperscript{14},
Federico Marini\textsuperscript{15},
Rita Giordano\textsuperscript{16},
Silvia Canelón\textsuperscript{17},
Anicet Ebou\textsuperscript{18},
Adithi Upadhya\textsuperscript{19},
Joselyn Chávez\textsuperscript{20},
Janani Ravi\textsuperscript{21*}
%,3\textcurrency}
\\
\bigskip
\textbf{1} Global Fishing Watch, Washington,  District of Columbia, United States of America
\\
\textbf{2} Instituto de Pesquisas Jardim Botânico do Rio de Janeiro, Rio de Janeiro, Rio de Janeiro, Brazil
\\
\textbf{3} International Institute for Sustainability, Rio de Janeiro, Rio de Janeiro, Brazil
\\
\textbf{4} Instituto Nacional de Tecnología Agropecuaria, Ruta Nac. N° 5, Km. 580, Anguil, La Pampa, Argentina; Universidad Nacional Guillermo Brown, Buenos Aires, Argentina
\\
\textbf{5} MetaDocencia, Argentina
\\
\textbf{6} Department of Statistics, University of Warwick, Coventry, UK; The R Foundation
\\
\textbf{7} Swiss Federal Institute for Forest, Snow and Landscape Research WSL, Birmensdorf, Switzerland
\\
\textbf{8} Instituto de Investigación e Ingeniería Ambiental, Universidad Nacional de San Martín - Consejo Nacional de Investigaciones Científicas y Técnicas, General San Martín, Buenos Aires, Argentina
\\
\textbf{9} KOF Swiss Economic Institute, ETH Zurich, Zurich, Switzerland
\\
\textbf{10} Institute of Systems, Molecular and Integrative Biology, University of Liverpool, Liverpool, L69 7ZB, UK; King Abdullah International Medical Research Center, Riyadh, Saudi Arabia
\\
\textbf{11} Dog Genetics LLC, Astoria, New York, United States of America
\\
\textbf{12} Instituto de Cálculo, FCEyN, UBA-CONICET, Buenos Aires, Argentina
\\
\textbf{13} Departamento de Botânica, Instituto de Biociências, Universidade de São Paulo, São Paulo, São Paulo, Brazil
\\
\textbf{14} Department of Statistics, University of California, Santa Cruz, California, United States of America
\\
\textbf{15} Institute of Medical Biostatistics, Epidemiology and Informatics (IMBEI), University Medical Center Mainz, Mainz, Germany
\\
\textbf{16} Royal Society of Chemistry, Thomas Graham House (290), Science Park, Milton Road, Cambridge, UK
\\
\textbf{17} Department of Biostatistics, Epidemiology \& Informatics, University of Pennsylvania, Philadelphia, Pennsylvania, United States of America.
\\
\textbf{18} Bioinformatics team, Département de Formation et de Recherches en Agriculture et Ressources Animales, Institut National Polytechnique Félix Houphouët-Boigny, BP 1313 Yamoussoukro, Côte d’Ivoire.
\\
\textbf{19} ILK Labs, Bengaluru, Karnataka, India
\\
\textbf{20} Instituto de Biotecnología, UNAM, Cuernavaca, Morelos, Mexico
\\
\textbf{21} Departments of Pathobiology and Diagnostic Investigation, Microbiology and Molecular Genetics, Michigan State University, East Lansing, Michigan, United States of America
\\
\bigskip

@@

% Use the asterisk to denote corresponding authorship and provide email address in note below.
*rocio.joo@globalfishingwatch.org, janani@msu.edu

@@

Conferences are spaces to meet and to network within and across academic and technical fields, to learn about new advances, and share our work. They can help define career paths and create long-lasting collaborations and opportunities. 
However, these opportunities are not equal for all. 
This article introduces ten simple rules to host an inclusive conference based on the authors' recent experience organizing the 2021 edition of the useR! statistical computing conference, which attracted a broad range of participants from academia, industry, government, and the non-profit sector. 
Coming from different backgrounds, career stages, and even continents, we embraced the challenge of organizing a high-quality virtual conference in the context of the COVID-19 pandemic and making it a kind, inclusive, and accessible experience for as many people as possible.
The rules result from our lessons learned before, during, and after the organization of the conference. 
They have been written mainly for potential organizers and selection committees of conferences and contain multiple practical tips to help a variety of events become more accessible and inclusive. We see this as a starting point for conversations and efforts towards building more inclusive conferences across the world.
* Translated versions of the English abstract and the list of rules are available in ten languages in S1 Text: Arabic, French, German, Italian, Japanese, Korean, Portuguese, Spanish, Tamil, and Thai.

@@

% Roadmap of the paper
Here, we present a set of ten rules based on the lessons we learned before, during, and after the organization of the useR! conference.
\linelabel{methods-a} The rules were first drafted by the core team and the members of the diversity, accessibility, and inclusion team of useR! 2021 who wrote their own ten simple rules (see S1 Text) based on the collective experience of organizing the conference and fueled by personal experiences and literature review.
A final list was proposed and then discussed and reviewed by the rest of the coauthors. \linelabel{methods-b}
The rules are organized in three sections: three foundation rules, six design rules, and a continuity rule (\textbf{Fig. \ref{fig:diagram}}).
The \textbf{foundation rules} comprise key elements to conceive the work on diversity and inclusion in any conference. 
\textbf{Rule 1} is about setting a vision of diversity and inclusion that should guide all the efforts and decision making in the organization.
\textbf{Rule 2} focuses on how to create a safe and welcoming environment for all the attendees. 
\textbf{Rule 3} highlights the importance of starting with an inclusive and diverse organizing team and provides tips on work dynamics.
The \textbf{design rules} focus on weaving inclusion into the conference design process.
In \textbf{Rule 4}, we introduce multiple ways to counteract bias in the conference program (keynotes, program committee, abstract selection, and thematic sessions). 
\textbf{Rule 5} provides advice for designing an inclusive online component in virtual and hybrid conferences.
\textbf{Rule 6} focuses on accessibility practices to include people with disabilities.
In \textbf{Rule 7}, we provide suggestions to account for linguistic diversity. 
\textbf{Rule 8} offers tips for developing an inclusive communication strategy. 
In \textbf{Rule 9}, we address budgeting for inclusive practices and helping participants with affordable registration costs, scholarships, and other forms of financial support.
Finally, 
\textbf{Rule 10}, the \textbf{continuity} rule, emphasizes the importance of self-assessment and advocates for making the conference part of a long-term commitment to inclusion and for passing the torch to future organizers.

@@

\section*{Acknowledgments}
The authors of this piece would like to thank every single member of the organizing team of useR! 2021 (\url{https://user2021.r-project.org/about/global-team}) for their valuable contribution to an inclusive conference experience and the R Foundation for trusting us with the organization of useR! 2021 and supporting us through the process. Special thanks to Saranjeet Kaur and gwynn sturdevant for suggestions to previous versions of the manuscript, Francisco Etchart for creating the Marmot figures, and Koki Tsuyuzaki, Kozo Nishida, KwangChun Lee, Arjun Krishnan, Kewalin Samart, Natnaree Yolnava, and Wanprakai Kaewphrae for their extended abstract translations. We would also like to thank Arjun Krishnan, Julia Ganz, Ingo Braasch, Susan Ewart, and Hilda Mejia-Abreu for their feedback on the manuscript.

@@

JR is funded by start-up award from Michigan State University (MSU). HT was supported by the Engineering and Physical Sciences Research Council [EPSRC EP/V052128/1] during the preparation of this article.

The funders had no role in study design, decision to publish, or preparation of the manuscript.

@@

\section*{Supporting Information Legends}

S1 Text - Abstract and list of rules in multiple languages
and rules proposed in the process of shaping these ten simple rules.
